---
title: "140XP Final Project Code"
author: "Sam Reade"
date: "2026-02-24"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading Necessary Libraries:

```{r}

library(ggplot2)
library(corrplot)
library(dplyr)

```

# Loading Dataset: 

```{r}

setwd("~/Desktop/Stats 140XP/Final Proj")

store_df <- read.csv("Sample - Superstore.csv")

```

# Investigating Data:

```{r}

head(store_df)
dim(store_df)
summary(store_df)

```

# Distribution of Profit and Sales:

```{r}

ggplot(store_df, aes(x = Profit)) +
  geom_histogram(bins = 40, fill = "purple") +
  theme_minimal()

ggplot(store_df, aes(x = Sales)) +
  geom_histogram(bins = 40, fill = "green") +
  theme_minimal()

```

# Profit By Category:

```{r}

ggplot(store_df, aes(x = Category, y = Profit)) +
  geom_boxplot() +
  theme_minimal()

```

# Profit By Region:

```{r}

ggplot(store_df, aes(x = Region, y = Profit)) +
  geom_boxplot() +
  theme_minimal()

```

# Correlation Matrix: 

```{r}

num_df <- store_df[, c("Sales", "Quantity", "Discount", "Profit")]

cor_matrix <- cor(num_df)
corrplot(cor_matrix, method = "color") 

```

# Data Wrangling:

Data wrangling = cleaning and transforming raw data into a usable format for analysis. 

```{r}

# Fixing Dates:
store_df <- store_df %>%
  mutate(
    Order.Date = as.Date(Order.Date, format = "%m/%d/%Y"),
    Ship.Date  = as.Date(Ship.Date, format = "%m/%d/%Y")
  )

# Time Variable: 
store_df <- store_df %>%
  mutate(
    Ship.Delay = as.numeric(Ship.Date - Order.Date),
    Order.Year = format(Order.Date, "%Y"),
    Order.Month = format(Order.Date, "%m")
  )

# Removing Useless Columns:
store_df <- store_df %>%
  select(-Row.ID, -Order.ID, -Customer.ID, -Product.ID)

# Creating Profit Margin Feature: 
store_df <- store_df %>%
  mutate(Profit.Margin = Profit / Sales)

# Creating Loss Indicator:
store_df <- store_df %>%
  mutate(Loss = ifelse(Profit < 0, 1, 0))

# Converting Categorical Data To Factors:
store_df <- store_df %>%
  mutate(
    Segment = as.factor(Segment),
    Region = as.factor(Region),
    Category = as.factor(Category),
    Sub.Category = as.factor(Sub.Category),
    Ship.Mode = as.factor(Ship.Mode)
  )

```

# Descriptive Stats: 

```{r}

# Summaries By Variable: 

summary(store_df[, c("Sales", "Profit", "Quantity", 
                     "Discount", "Ship.Delay", "Profit.Margin")])

```

```{r}

# Grouped Statistics By Region: 

store_df %>%
  group_by(Region) %>%
  summarise(
    Mean_Profit = mean(Profit),
    Mean_Sales = mean(Sales),
    Loss_Rate = mean(Loss)
  )

# Loss Rate By Discount: 

store_df %>%
  group_by(Discount) %>%
  summarise(Loss_Rate = mean(Loss))

```

# Checking Skewness Statistically: 

```{r}

library(e1071)

skewness(store_df$Sales)
skewness(store_df$Profit)

```

# Transforming Sales and Profit: 

```{r}

store_df <- store_df %>%
  mutate(log_Sales = log(Sales))

skewness(store_df$log_Sales)

store_df <- store_df %>%
  mutate(log_Profit = log(abs(Profit) + 1))

skewness(store_df$log_Profit)

```

# Distributions After Transformations: 

```{r}

ggplot(store_df, aes(x = log_Sales)) +
  geom_histogram(bins = 40, fill = "darkgreen") +
  theme_minimal() +
  labs(title = "Distribution of Log(Sales)")

ggplot(store_df, aes(x = log_Profit)) +
  geom_histogram(bins = 40, fill = "orange") +
  theme_minimal() +
  labs(title = "Distribution of Log(|Profit| + 1)")

```

# Profit VS. Discount: 

```{r}

ggplot(store_df, aes(x = Discount, y = Profit)) +
  geom_point(alpha = 0.2) +
  theme_minimal()

```

# Checking Correlation: 

```{r}

cor(store_df[, c("Sales", "Profit", "Discount", "Quantity")])

```

# Checking Assumptions: 

```{r}

shapiro.test(sample(store_df$Profit, 500))

```


# Hypothesis Testing:

# Do Discounts Reduce Profit?

```{r}

no_discount <- store_df$Profit[store_df$Discount == 0]
discounted <- store_df$Profit[store_df$Discount > 0]

t.test(no_discount, discounted, alternative = "greater")

```

# Is Average Profit Different Across Regions?

```{r}

anova_model <- aov(Profit ~ Region, data = store_df)
summary(anova_model)

TukeyHSD(anova_model)

```

# Is Discount Negatively Correlated with Profit?

```{r}

cor.test(store_df$Discount, store_df$Profit, alternative = "less")

```











